SHIV: Reducing Supervisor Burden using Support Vectors for Efficient Learning from Demonstrations in High Dimensional State Spaces

### 2015-09-16 paper.tex
Updates from ShareLaTeX
Anca's edits

In model-free robot Learning from Demonstration (LfD), a robot learns to perform a task, such as driving or grasping an object in a cluttered environment, from examples provided by a  supervisor, usually a human.  In such problems, the robot does not have access to either the cost function that it should optimize, nor the dynamics model. The former occurs when it is difficult to specify how to trade-off various aspects that matter, like a car trying to drive on the road and dodge other cars \cite{abbeel2004apprenticeship}.  The latter occurs when either the system or the interaction with the world is difficult to characterize like when a robot is trying to grasp an object in clutter and doesn't have an accurate model of the objects dynamics.

Rather than explicitly learning the cost function, like in Inverse Reinforcement Learning \cite{ng2000algorithms}, and the dynamics model, and then using optimization to produce a policy for the task, in model-free LfD the robot learns the policy directly from supervisor examples mapping states to controls~\cite{argall2009survey}. Learning from demonstration has been used successfully in recent year for a large number of robotic tasks, including helicopter maneuvering \cite{abbeel2007application},car parking \cite{abbeel2008apprenticeship},  and robotic surgery \cite{van2010superhuman}.

@@

We present an algorithm for online LfD that can actively decide whether it is necessary to ask the supervisor to label the new states that the robot encounters. SHIV  (Svm-based reduction in Human InterVention) reduces supervisor burden by only requesting supervision for risky states. SHIV uses stream-based active learning with a query selection method that evaluates risk in a manner tailored to the non-stationary state distributions the robot encounters as it iteratively executes learned policies. We build on the One Class SVM method for approximating quantile level sets \cite{scholkopf2001estimating} to provide a classification method for deciding if a state is risky or safe.

@@

\noindent\textbf{Active Learning from Demonstration}
We formulate applying active learning to Online Lfd as a stream-based active learning problem ~\cite{atlas1990training,cohn1994improving}. In stream based active learning, the decision of whether to query the supervisor (for a label or control signal) or not is not over the entire state space (like in traditional pool-based active learning), but on states drawn one at a time from some data stream. In  online LfD, this data stream is the states encountered when the robot is executing the current best policy.

Aside from the data stream, stream-based active learning has another ingredient: a query selection method. A query selection method is deciding whether or not to query. Typical query selection methods are estimator-centric, evaluating risk using the estimator, e.g. distance from the classification hyperplane \cite{tong2002support}, as in Fig. \ref{fig:support_example}(a); or query by committee  \cite{breiman1996bagging}, which uses a committee of hypothesized estimators that are trained on different subsets of the training data. Risk in query by committee is based on the level of agreement or disagreement among these hypotheses, with higher levels of disagreement for a state leading to higher chances of querying that state (Fig. \ref{fig:support_example}(b)).

@@

The problem lies in being estimator-centric. The  estimator is a function of the current distribution. Therefore, it no longer provides a good measure of confidence when that distribution changes, as it does when the robot is rolling out a learned policy and starts encountering further away states. Instead of relying on the estimator, our measure of risk explicitly identifies when states are drawn from a different distribution, i.e. when they are novel.

@@

Our work builds on Kim et al. \cite{kim2013maximum} who apply novelty detection for safe online learning from demonstrations to enable a supervisor to take over and prevent unsafe states.  In their MMD-LI algorithm, risk is evaluated with Maximal Mean Discrepancy, which for a single point is the sum of a kernel density estimate and variance on the dataset, which works well in low dimensional state spaces.  MMD-LI also reduces supervisor burden because the supervisor only takes control and provides demonstrations in risky states. However, having the supervisor take over can bias the states sampled, which can in practice lead to a loss in performance \cite{ross2010reduction}.  In contrast, we focus on active learning for online LfD for high-dimensional state spaces, do not assume the supervisor is always available, and propose an alternative to non-parametric density estimation which can require data exponential in the dimension of the state-space
\cite{nadaraya1964estimating}.

An alternative to evaluating the probability of a data point is to measure distance to its nearest neighbors \cite{knox1998algorithms}. However, this approach was shown to be
susceptible to issues, since nearest neighbors  incorporates only local information about the data. For example, a group of outliers can be close together, but significantly far from the majority of the data and nearest neighbors would mark them as not novel \cite{hodge2004survey}.

Our approach is based on the One Class SVM proposed by Scholk{\"o}pf et al., which estimates a particular quantile level set for the training data by solving a convex quadratic program  to the find support
vectors~\cite{scholkopf2001estimating}. The method has been theoretically shown to approximate the quantile levelset of a density estimate
asymptotically for correctly chosen bandwidth settings and in the case of a normalized Gaussian kernel function \cite{vert2006consistency}. 
In \cite{liu2014unsupervised}, the One Class SVM has furthermore been used for novelty detection in high-dimensional image data. To our knowledge, incorporating a query selection method for nonstationary distributions, such as the One Class SVM, into online LfD is a novel contribution.

@@

The decision function parametrization consists of the kernel bandwidth $\sigma$ in $g_s$. We treat $\sigma$ as a "risk sensitivity" parameter. For two reasons: 1)The expected number of examples, after a policy roll out, the supervisor can be asked is  $T*\int_\bx \mathbf{1}(g_{\sigma}(\bx) == 0) p(\bx|\theta)d\bx$. Thus, smaller $\sigma$ corresponds to asking for more examples. 2) A relation exists between how smooth the supervisor's policy, $\tilde{\pi}$ and how many examples are needed to learn it. Thus, a large $\sigma$ can be dangerous for policies with sharp variation because it will treat points as safe that are really risky.

@@

\section{Experiments}

@@

We then  compare our query selection method with those typically used in active learning. We show that for a non-stationary state distribution like ours, the notion of risk based on novelty and misclassified regions performs better than confidence, query-by-committee based methods and the One Class SVM with out the hole carving out modification. We continue with a sensitivity analysis, which suggests that the performance of SHIV is robust to the choice of how risky the robot is allowed to be (the $\sigma$ parameter from Eq. \ref{eq:decision_func}).

@@

 We compare four active learning methods. We compare our combined notion of risk (Fig. \ref{fig:support_example}(d)), with risk based on novelty alone (Fig. \ref{fig:support_example}(c)) in order to test whether carving out regions that have been misclassified previously is valuable.

@@

\section{Discussions and Future Work}
Online LfD have been successful on array of tasks~\cite{NIPS2014_5421,ross2010reduction}, however can place significant burden on a supervisor. Our algorithm, SHIV, implements stream-based active learning with a query selection method that evaluates risk in a manner tailored to non-stationary state distributions. We empirically evaluated our method on three different domains: a driving simulator, grasping in clutter and surgical needle insertion. Results suggests up to a $70\%$ reduction in the number of queries to the supervisor.

>>>>

In model-free robot Learning from Demonstration (LfD), a robot learns to perform a task, such as driving or grasping an object in a cluttered environment, from examples provided by a  supervisor, usually a human.  In such problems, the robot does not have access to either the cost function that it should optimize, nor the dynamics model. The former occurs when it is difficult to specify how to trade-off various aspects that matter, like a car trying to drive on the road and dodge other cars \cite{abbeel2004apprenticeship}.  The latter occurs when either the system or the interaction with the world is difficult to characterize, like when a robot is trying to grasp an object in clutter and does not have an accurate model of the objects dynamics.

Rather than explicitly learning the cost function (like in Inverse Reinforcement Learning \cite{ng2000algorithms}) and the dynamics model, and then using optimization to produce a policy for the task, in model-free LfD the robot learns the policy directly from supervisor examples, mapping states to controls~\cite{argall2009survey}. Learning from demonstration has been used successfully in recent year for a large number of robotic tasks, including helicopter maneuvering \cite{abbeel2007application},car parking \cite{abbeel2008apprenticeship},  and robotic surgery \cite{van2010superhuman}.

@@

We present an algorithm for online LfD that can actively decide whether it is necessary to ask the supervisor to label the new states that the robot encounters. SHIV  (Svm-based reduction in Human InterVention) reduces supervisor burden by only requesting supervision for risky states. SHIV uses stream-based active learning with a query selection method that evaluates risk in a manner tailored to the non-stationary state distributions the robot encounters as it iteratively executes learned policies, and tailored to high dimensional state spaces. We build on the One Class SVM method for approximating quantile level sets \cite{scholkopf2001estimating} to provide a classification method for deciding if a state is risky or safe.

@@

\noindent\textbf{Active Learning from Demonstration}
We formulate reducing supervisor burden in nnline Lfd as a stream-based active learning problem ~\cite{atlas1990training,cohn1994improving}. In stream based active learning, the decision of whether to query the supervisor (for a label or control signal) or not is not over the entire state space (like in traditional pool-based active learning), but on states drawn one at a time from some data stream. In  online LfD, this data stream is the states encountered when the robot is executing the current best policy.

Aside from the data stream, stream-based active learning has another ingredient: a query selection method, deciding whether or not to query. Typical query selection methods are estimator-centric, evaluating risk using the estimator, e.g. distance from the classification hyperplane \cite{tong2002support}, as in Fig. \ref{fig:support_example}(a); or query by committee  \cite{breiman1996bagging}, which uses a committee of hypothesized estimators that are trained on different subsets of the training data. Risk in query by committee is based on the level of agreement or disagreement among these hypotheses, with higher levels of disagreement for a state leading to higher chances of querying that state (Fig. \ref{fig:support_example}(b)).

@@

The problem lies in being estimator-centric. The  estimator is a function of the current distribution. Therefore, it no longer provides a good measure of confidence when that distribution changes. And the distribution does change when the robot is rolling out a learned policy and starts encountering further away states. Instead of relying on the estimator, our measure of risk explicitly identifies when states are drawn from a different distribution, i.e. when they are novel. Our experiments in Section \ref{sec:experiments} suggest that our measure of risk is more accurate in online LfD problems than estimator-centric measures.

@@

Our work builds on Kim et al. \cite{kim2013maximum} who apply novelty detection to safe online LfD to enable a supervisor to take over and prevent unsafe states.  In their MMD-LI algorithm, risk is evaluated with Maximal Mean Discrepancy. For a single state, this is the sum of a kernel density estimate and the variance of the dataset, which works well in low dimensional state spaces.  MMD-LI also reduces supervisor burden because the supervisor only takes over in risky states, as opposed to providing labels for all new states. 
In contrast, we focus on active learning for online LfD in high-dimensional state spaces. Unlike safe learning, our problem does not require the supervisor to take over. This enables us to avoid the supervisor biasing the states sampled (which can in practice lead to a loss in performance \cite{ross2010reduction}), and also avoid assuming the supervisor is always available. On the other hand, our problem does require novelty detection in high dimensional state spaces, where non-parametric density estimation does not scale very well: it can require data exponential in the dimension of the state space \cite{nadaraya1964estimating}. Our results (Section \ref{sec:experiments}) suggests that in our high dimensional state spaces, SHIV achieves an error rate half of that of kernel density estimate-bases techniques like MMD.
%However, having the supervisor take over can bias the states sampled, which can in practice lead to a loss in performance \cite{ross2010reduction}. In contrast, we focus on active learning for online LfD for high-dimensional state spaces, do not assume the supervisor is always available, and propose an alternative to non-parametric density estimation which can require data exponential in the dimension of the state-space \cite{nadaraya1964estimating}.

An alternative to kernel density estimates is to measure distance to its nearest neighbors \cite{knox1998algorithms}. However, this approach was shown to be
susceptible to issues, since nearest neighbors  incorporates only local information about the data. For example, a group of outliers can be close together, but significantly far from the majority of the data and nearest neighbors would mark them as not novel \cite{hodge2004survey}.

Our approach is based on the One Class SVM proposed by Scholk{\"o}pf et al., which estimates a particular quantile level set for the training data by solving a convex quadratic program  to the find support
vectors~\cite{scholkopf2001estimating}. The method has been theoretically shown to approximate the quantile levelset of a density estimate
asymptotically for correctly chosen bandwidth settings and in the case of a normalized Gaussian kernel function \cite{vert2006consistency}. 
In \cite{liu2014unsupervised}, the One Class SVM has furthermore been used for novelty detection in high-dimensional image data. To our knowledge, incorporating a query selection method for high-dimensional spaces and nonstationary distributions, such as the One Class SVM, into online LfD is a novel contribution.

@@

The decision function parametrization consists of the kernel bandwidth $\sigma$ in $g_s$. We treat $\sigma$ as a "risk sensitivity" parameter (and study its implications in Section \ref{sec:experiments}). For two reasons: 1)The expected number of examples, after a policy roll out, the supervisor can be asked is  $T*\int_\bx \mathbf{1}(g_{\sigma}(\bx) == 0) p(\bx|\theta)d\bx$. Thus, smaller $\sigma$ corresponds to asking for more examples. 2) A relation exists between how smooth the supervisor's policy, $\tilde{\pi}$ and how many examples are needed to learn it. Thus, a large $\sigma$ can be dangerous for policies with sharp variation because it will treat points as safe that are really risky.

@@

\section{Experiments}\label{sec:experiments}

@@

We then  compare our query selection method with those typically used in active learning. We show that for a non-stationary state distribution like ours, the notion of risk based on novelty and misclassified regions performs better than confidence, query-by-committee based methods, Maximum Mean Discrepancy, and the One Class SVM without the carving out misclassified holes modification. We continue with a sensitivity analysis, which suggests that the performance of SHIV is robust to the choice of how risky the robot is allowed to be (the $\sigma$ parameter from Eq. \ref{eq:decision_func}).

@@

 We compare five active learning methods. We compare our combined notion of risk (Fig. \ref{fig:support_example}(d)), with risk based on novelty alone (Fig. \ref{fig:support_example}(c)) in order to test whether carving out regions that have been misclassified previously is valuable.

@@

\section{Discussions and Future Work}
Online LfD has been successful on an array of tasks~\cite{NIPS2014_5421,ross2010reduction}, however it can place significant burden on a supervisor. Our algorithm, SHIV, implements stream-based active learning with a query selection method that evaluates risk in a manner tailored to non-stationary and high-dimensional state distributions. We empirically evaluated our method on three different domains: a driving simulator, grasping in clutter and surgical needle insertion. Results suggests up to a $70\%$ reduction in the number of queries to the supervisor.
